openTSNE: a modular \proglang{Python} library for t-SNE dimensionality reduction and embedding

### 2023-05-23 2023-05-revision-1/main.tex
paper: ensure that uniform kernel doesn't come out of nowhere

\begin{enumerate}
\item The t-SNE algorithm requires a specification of an affinity model to describe the similarities between data points. Standard t-SNE predominantly deals with tabular data and computes similarities using Equation~\ref{eq:p_ij}. However, \citet{kobak2019art} showed that using mixtures of Gaussians instead of a single Gaussian often reveals better global consistency of the resulting visualization (albeit at some computational cost). Alternatively, one can speed up the overall computation by choosing a uniform kernel, which yields quantitatively similar results to standard t-SNE. \code{openTSNE.affinity} provides several affinity models to choose from but defaults to the standard perplexity-based model proposed in the original publication.

In \opentsne, it is straightforward to define and use a custom affinity model. For instance, one can extend the \code{openTSNE.affinity.Affinities} class and provide an affinity model for relational data using a normalized adjacency matrix. Such extension enables rapid experimentation and opens the doors to the visualization of non-tabular data.

>>>>

\begin{enumerate}
\item The t-SNE algorithm requires the specification of an affinity model to describe the similarities between data points $p_{ij}$. Standard t-SNE uses perplexity-based Gaussian kernels with varying bandwidths as specified in Equation~\ref{eq:p_ij}. However, other affinity models can also be used to highlight different aspects of the data or visualize other data modalities. For instance, \citet{kobak2019art} showed that using mixtures of Gaussians often better reveals global cluster organization in the resulting visualization (albeit at some computational cost). Alternatively, one can use a uniform affinity model that assigns equal similarities $p_{ij}$ to a pre-specified number of nearest neighbors. This yields quantitatively similar results to standard t-SNE at a lower computational cost, making it particularly suitable for larger data sets. Visualizing different data modalities, such as graphs, can also be achieved by providing a precomputed distance matrix that might consist of, for instance, the shortest paths between nodes. The \code{openTSNE.affinity} module provides a selection of commonly used affinity models but defaults to the standard perplexity-based Gaussian model, as proposed in the original publication. Users can also define custom affinity models by extending the \code{openTSNE.affinity.Affinities} class. The resulting class can then be used as a drop-in replacement for the built-in affinity models.
